### 2020-09-24 LaTeX/sections/abstract.tex
final edits

\begin{abstract}
Order-driven exchanges on Ethereum tend to operate in a fully centralized fashion, or with some functionality performed on-chain (\eg loading accounts, order cancelation) and some off-chain functionality (\eg matching orders). Fully on-chain solutions have a superior threat model and were experimented with in the early days of Ethereum, but they have been largely abandoned for performance reasons. As block\-chain performance continues to improve (both in research and in practice) and on-chain dealer-based exchanges like Uniswap (an alternative to order-based exchanges) rise in popularity, it seems to be an ideal time to revisit the idea of a fully on-chain order-driven exchange and benchmark what is possible performance-wise. We design and optimize an Ethereum-based call market exchange, \cm, and conduct a variety of experiments to show the current benchmark is around one hundred trade executions per block. We also explore numerous extensions for further improvements, including a variant using off-chain execution (`rollups').
\end{abstract}

>>>>

\begin{abstract}
Order-driven exchanges on Ethereum tend to operate in a fully centralized fashion, or with some functionality performed on-chain (\eg loading accounts, order cancellation) and some off-chain functionality (\eg matching orders). Fully on-chain solutions have a superior threat model and were experimented with in the early days of Ethereum, but they have been largely abandoned for performance reasons. As block\-chain performance continues to improve (both in research and in practice) and on-chain dealer-based exchanges like Uniswap (an alternative to order-based exchanges) rise in popularity, it seems to be an ideal time to revisit the idea of a fully on-chain order-driven exchange and benchmark what is possible performance-wise. We design and optimize an Ethereum-based call market exchange, \cm, and conduct a variety of experiments to show the current benchmark is around one hundred trade executions per block. We also explore numerous extensions for further improvements, including a variant using off-chain execution (`rollups').
\end{abstract}

### 2020-09-24 LaTeX/sections/paper.tex
final edits

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, processing all the trades and closing the market is the most time consuming step. The most critical design decision is the data structure for holding orders. While data structures are well studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object oriented programming) that create difficulties in assessing which will perform best without actually deploying and evaluating each variant.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{fig/random_insertion2.png}
\caption{\footnotesize{Gas costs for enqueuing 50 random integers into five priority queue variants. For the x-axis, a value of 9 indicates it is the 9th number entered in the priority queue. The y-axis is the cost of enqueuing in gas.}  \label{fig:random_insertion}}
\end{figure}

@@

The main research question is how many orders can be processed in a single Ethereum transaction when closing the call market, using Ethereum today. As our previous experiments indicated, the choice of priority queue (PQ) implementation is the main influence on performance (see Table~\ref{tab:worst_case_matching}). We implemented a generic call market that interfaces with a generic PQ (at its own contract address) and ran experiments for each PQ implementation. We looked at the \textit{worst-case} for performance which is when every submitted bid and ask is marketable (\ie will require fulfillment). In the first two columns, we determined the highest number of orders that can be processed in a single call to the \texttt{closeMarket()} and not exceed the the current Ethereum block gas limit of 11,741,495. Since Ethereum will become more efficient over time, we also were interested in how much gas it would cost to execute 1000 pairs of orders which is given in the third column. The fourth column indicates the cost of submitting a bid or ask --- since this costs will vary depending on how many orders are already submitted (recall Figure~\ref{fig:random_insertion}), we average the cost of 200 order submissions.

As expected, the numbers closely track the performance of the PQ itself suggesting the PQ is indeed the main influence on performance. In Ethereum today, call markets appear to be limited to processing about a hundred orders per transaction. If markets open on every other block and the call market could monopolize an entire block to close, a few hundred orders per minute (worst-case) can be processed. The main takeaway is that the transparency, front-running resistance, and low barrier to entry of Ethereum comes with an enormous cost (\ie an institutional exchange like NASDAQ can process 100K trades per second). That said, many exchanges trade the same assets under different trading rules (\ie market fragmentation) because traders have different preferences. \cm can work today in some circumstances like very low liquidity tokens, or markets with high volumes and a small number of traders (\eg liquidation auctions).

@@

As a module, we try to keep \cm as simple as possible, and allow customizations and extensions as appropriate. We discuss potential modifications here.

@@

We implement a very simple market where orders all have the same volume: 1 token. Traders submit multiple orders to increase volume. To extend orders with a volume field, it must first be determined if the token is non-divisible or divisible (and to what precision). In this case, \texttt{closeMarket()} needs to look at multiple bids to fulfill a single ask (and vice-versa) and declare its trading rules on partial fills. We also simplify the market rules around ties on price: we execute them in a FIFO manner (breaking front-running resistance for ties on prices). A common trading rule (that does resist front-running) is to fill ties in proportion to their volume (\ie \textit{pro rata} allocation)\footnote{If Alice and Bob bid the same price for 100 tokens and 20 tokens respectively, and there are only 60 tokens left in marketable asks, Alice receives 50 and Bob 10.} however this approach is also contingent on the divisibility of tokens. Consider the following corner-case: 3 equally priced bids of 1 non-divisible token and 1 ask at the same price. There is no good option: (1) the bids could all be dropped (fair but not market efficient), (2) bids could be prioritized based on time as in \cm (front-running is viable, but in this corner case only), or (3) the bid could be randomly chosen (\cf Libra~\cite{mavroudis2019libra}); blockchain `randomness' is generally deterministic and manipulatable by miners~\cite{bonneau2015random,buenz2017proofs} and counter-measures could take a few blocks to select~\cite{boneh2018verifiable}).

@@

If the call market runs as a DApp on Ethereum, it seems difficult to open and close the market every block. Someone needs to call \texttt{closeMarket()} for every block (we return to who this is next) but the market will only work as intended if miners execute this function after every \texttt{submitBid()} and \texttt{submitAsk()} invocation. Since price improvements are paid to the miners, the miner is incentivized to run \texttt{closeMarket()} last to make the most profit. Perhaps in a world where these markets became immensely popular, miners would actually do this, but as it stands, miners do not generally search for the transaction ordering that nets them the most profit.

@@

The final issue is who invokes \texttt{closeMarket()} every other block? There are actually two issues here: the issue of scheduling the function call and the issue of paying for it. For scheduling the function call, we can do one of the following: rely on market participants who are eager to trade to reopen the market, offer a bounty to reopen the market, or use an external service like Ethereum Alarm Clock.\footnote{\url{https://ethereum-alarm-clock-service.readthedocs.io/}} Next we consider the second issue of who pays to close the market.

@@

It is unlikely that a mining software will become aware of specific DApps, but miners are paid all price improvements in the market and so it is possible that running \texttt{closeMarket()} would pay for itself. However we consider two other (more realistic) scenarios. One solution requires a modified closing function, \texttt{closeMarket(n)}, that only processes $n$ orders at a time (until there is no orders left in the list). Once the time window for submitting orders is past, a new market is created (without settling the previous market). Every order submission on the new market also requires to run, say, \texttt{closeMarket(10)} on the older market, thus progressively closing the previous market while accepting orders to the new market. This solution pattern has two issues: first, amortizing the cost of closing the market amongst the early traders of the new market is an added incentive to not submit orders early to the market; the second related issue is if not enough traders submit orders in the new market, the old market never closes (resulting in a backlog of old markets waiting to close).

>>>>

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, closing the market and processing all the orders is the most time consuming step. The most critical design decision is the data structure for holding orders. While data structures are well studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object oriented programming) that create difficulties in assessing which will perform best without actually deploying and evaluating each variant.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{fig/random_insertion2.png}
\caption{\footnotesize{Gas costs for enqueuing 50 random integers into five priority queue variants. For the x-axis, a value of 9 indicates it is the 9th integer entered in the priority queue. The y-axis is the cost of enqueuing in gas.}  \label{fig:random_insertion}}
\end{figure}

@@

The main research question is how many orders can be processed in a single Ethereum transaction when closing the call market, using Ethereum today. As our previous experiments indicated, the choice of priority queue (PQ) implementation is the main influence on performance (see Table~\ref{tab:worst_case_matching}). We implemented a generic call market that interfaces with a generic PQ (at its own contract address) and ran experiments for each PQ implementation. We looked at the \textit{worst-case} for performance which is when every submitted bid and ask is marketable (\ie will require fulfillment). In the first two columns, we determine the highest number of orders that can be processed in a single call to the \texttt{closeMarket()} and not exceed the the current Ethereum block gas limit of 11,741,495. Since Ethereum will become more efficient over time, we also were interested in how much gas it would cost to execute 1000 pairs of orders which is given in the third column. The fourth column indicates the cost of submitting a bid or ask --- since this costs will vary depending on how many orders are already submitted (recall Figure~\ref{fig:random_insertion}), we average the cost of 200 order submissions.

As expected, the numbers closely track the performance of the PQ itself suggesting the PQ is indeed the main influence on performance. In Ethereum today, call markets appear to be limited to processing about a hundred orders per transaction. If markets open on every other block and the call market could monopolize an entire block to close, a few hundred orders per minute (worst-case) can be processed. The main takeaway is that the transparency, front-running resistance, and low barrier to entry of Ethereum come with an enormous cost (\ie an institutional exchange like NASDAQ can process 100K trades per second). That said, many exchanges trade the same assets under different trading rules (\ie market fragmentation) because traders have different preferences. \cm can work today in some circumstances like very low liquidity tokens, or markets with high volumes and a small number of traders (\eg liquidation auctions).

@@

As a module, we tried to keep \cm as simple as possible, and allow customizations and extensions as appropriate. We discuss potential modifications here.

@@

We implemented a very simple market where orders all have the same volume: 1 token. Traders submit multiple orders to increase volume. To extend orders with a volume field, it must first be determined if the token is non-divisible or divisible (and to what precision). In this case, \texttt{closeMarket()} needs to look at multiple bids to fulfill a single ask (and vice-versa) and declare its trading rules on partial fills. We also simplify the market rules around ties on price: we execute them in a FIFO manner (breaking front-running resistance for ties on prices). A common trading rule (that does resist front-running) is to fill ties in proportion to their volume (\ie \textit{pro rata} allocation)\footnote{If Alice and Bob bid the same price for 100 tokens and 20 tokens respectively, and there are only 60 tokens left in marketable asks, Alice receives 50 and Bob 10.} however this approach is also contingent on the divisibility of tokens. Consider the following corner-case: 3 equally priced bids of 1 non-divisible token and 1 ask at the same price. There is no good option: (1) the bids could all be dropped (fair but not market efficient), (2) bids could be prioritized based on time as in \cm (front-running is viable, but in this corner case only), or (3) the bid could be randomly chosen (\cf Libra~\cite{mavroudis2019libra}); blockchain `randomness' is generally deterministic and manipulatable by miners~\cite{bonneau2015random,buenz2017proofs} and counter-measures could take a few blocks to select~\cite{boneh2018verifiable}).

@@

If the call market runs as a DApp on Ethereum, it seems difficult to open and close the market every block. Someone needs to call the \texttt{closeMarket()} for every block (we return to who this is next) but the market will only work as intended if miners execute this function after every \texttt{submitBid()} and \texttt{submitAsk()} invocation. Since price improvements are paid to the miners, the miner is incentivized to run \texttt{closeMarket()} last to make the most profit. Perhaps in a world where these markets became immensely popular, miners would actually do this, but as it stands, miners do not generally search for the transaction ordering that nets them the most profit.

@@

The final issue is who invokes \texttt{closeMarket()} every other block? There are actually two issues here: the issue of scheduling the function call and the issue of paying for it. For scheduling the function call, we can do one of the following: rely on market participants, who are eager to trade, to reopen the market, offer a bounty to reopen the market, or use an external service like Ethereum Alarm Clock.\footnote{\url{https://ethereum-alarm-clock-service.readthedocs.io/}} Next we consider the second issue of who pays to close the market.

@@

It is unlikely that a mining software will become aware of specific DApps, but miners are paid all price improvements in the market and so it is possible that running \texttt{closeMarket()} would pay for itself. However we consider two other (more realistic) scenarios. One solution requires a modified closing function, \texttt{closeMarket(n)}, that only processes $n$ orders at a time (until the order book is empty). Once the time window for submitting orders is past, a new market is created (without settling the previous market). Every order submission on the new market also requires to run, say, \texttt{closeMarket(10)} on the older market, thus progressively closing the previous market while accepting orders to the new market. This solution pattern has two issues: first, amortizing the cost of closing the market amongst the early traders of the new market is an added incentive to not submit orders early to the market; the second related issue is if not enough traders submit orders in the new market, the old market never closes (resulting in a backlog of old markets waiting to close).

### 2020-09-24 LaTeX/tables/functions.tex
full round edit (#2)
final edits

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	depositToken()                           & Deposits ERC20 standard compliant tokens in \cm smart contract \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	depositEther()                           	& Deposits ETH in \cm smart contract                        			 \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	openMarket()                             & Opens the market                                                    					 \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	closeMarket()                            & Closes the market                                                    					\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	submitBid()                              & Inserts the upcoming bid order messages inside the priority queue    \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	submitAsk()                              & Inserts the upcoming ask order messages inside the priority queue    \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	matchOrders()                           & Matches the orders against each other                                			\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	claimTokens()                           & Transfers collateralized tokens back to the trader                       			\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	claimEther()                             & Transfers collateralized ETH back to the trader                       			\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
\end{tabular}
\caption{\footnotesize{Primary operations of \cm smart contract.}
\label{tab:cm_functions}}
\end{table}

>>>>

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	depositToken()                           & Deposits ERC20 standard compliant tokens in \cm smart contract \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	depositEther()                           	& Deposits ETH in \cm smart contract                        			 \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	openMarket()                             & Opens the market                                                    					 \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	closeMarket()                            & Closes the market and processes the orders       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	submitBid()                              & Inserts the upcoming bid order messages inside the priority queue    \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	submitAsk()                              & Inserts the upcoming ask order messages inside the priority queue    \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	claimTokens()                           & Transfers tokens to the traders                       			\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
	claimEther()                             & Transfers ETH to the traders                       			\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
\end{tabular}
\caption{\footnotesize{Primary operations of \cm smart contract.}
\label{tab:cm_functions}}
\end{table}
